Simulating International Energy Security

### 2013-12-08 Writeups/CSS739 Paper/Paper.tex
Paper updates

I run the model for 100 iterations for each permutation of the Contagion and Assistance flags, 400 iterations in total; each iteration was run for 60 ticks, or 5 simulated years. I collect the country-level and global indicators for each month of each iteration.

Figure X shows a histogram of the Global Overall Ratio (the ratio between secure demand and supply) across all iteration-months. It appears to approximate a skew normal distribution, centered on 1 (a balance between supply and demand) but with a longer tail to the right -- supply shocks are more likely than demand shocks, and the most extreme supply shocks the model generates, though low-probability, are more extreme than any demand shocks.

@@

Upon visual examination, the distributions with Contagion set to 'On' and 'Off' appear extremely similar. I conduct two statistical tests to determine whether they are in fact drawn from the same underlying distribution -- that is, whether there is no actual difference between them. Both the Kolmogorov-Smirnov  and Mann-Whitney tests indicate that the distributions are in fact different. Nevertheless, the differences are small. The mean ratio for the No Contagion scenario is 1.046, compared to 1.068 for the Contagion scenario. Both a parametric T-Test and nonparametric Wilcoxon Rank-Sum confirm that these means are indeed different, though the difference is small. Similarly, the variance for the No Contagion scenario is 0.045, while the Contagion variance is 0.051. Thus, conflict contagion pushes the scenarios away from the mean, yielding a wider range of outcomes.

I repeat the same analysis for the the Assistance disabled and enabled scenarios. Again, while the distributions appear extremely similar, both the Kolmogorov-Smirnov and Mann-Whitney tests show that they are different. The mean and variance for Assistance Off are 1.062 and 0.05, respectively, compared to 1.053 and 0.046 when Assistance is enabled. This confirms with our intuition -- international assistance reduces the uncertainty of energy availability, and increases overall energy security, albeit by a small number.

@@

Figure XX shows a map of the overall risk of a major supply crisis, with the supply ratio rising above 2. Several things immediately jump out: several countries at the highest risk countries are major net oil exporters, including Canada, Mexico and Iran. While these countries derive a large fraction of their consumption from domestic production, they also rely on imports from abroad

@@

>>>>

I run the model for 250 iterations for each permutation of the Contagion and Assistance flags, 1,000 iterations in total; each iteration was run for 60 ticks, or 5 simulated years. I collect the country-level and global indicators for each month of each iteration.

Figure X shows a histogram of the Global Overall Ratio (the ratio between secure demand and supply) across all iteration-months. It appears to approximate a skew normal distribution, centered very close to 1 (a balance between supply and demand) but with a longer tail to the right -- supply shocks are more likely than demand shocks, and the most extreme supply shocks the model generates, though low-probability, are more extreme than any demand shocks.

@@

Upon visual examination, the distributions with Contagion set to 'On' and 'Off' appear extremely similar. I conduct several statistical tests to determine whether they are in fact drawn from the same underlying distribution -- that is, whether there is no actual difference between them. Both the Kolmogorov-Smirnov  and Mann-Whitney tests indicate that the distributions are in fact different. Nevertheless, the differences are small. The mean ratio for the No Contagion scenario is 1.037, compared to 1.049 for the Contagion scenario. Both a parametric T-Test and nonparametric Wilcoxon Rank-Sum confirm that these means are indeed different, and neither the traditional (frequentist) not Bayesian confidence intervals overlap. Similarly, the variance for the No Contagion scenario is 0.025, while the Contagion variance is 0.03. Thus, conflict contagion pushes the scenarios away from the mean, yielding a wider range of outcomes.

I repeat the same analysis for the the Assistance disabled and enabled scenarios. However, in this case, the statistical tests cannot reject the hypothesis that the resulting distributions are the same, and the means are within each other's confidence intervals.

@@

Figure XX shows a map of the overall risk of a major supply crisis, with the supply ratio rising above 1.5 (that is, demand exceeds secure supply by at least 50\%). Several things immediately jump out: major oil producers appear to be at the least risk, as their consumption of domestic production reduces their exposure to import disruption. Russia stands out as the one major producer to face a significant supply risk. The countries of the so-called `Global South' appears to face higher risks than the more developed world, likely indicating a lower diversity of import sources.

Figure XX shows a similar map for demand risk, and indicates the likelihood of total possible exports exceed secure exports by at least 50\%. While there are fewer countires with very low risk, there are fewer with extremely high risk as well. Interestingly, the major Middle East producers appear to face relatively low demand risk, likely indicating a robust set of trade relationships with realively stable countries. In contrast, the major non-US producers in the Americas -- Canada, Mexico and Venezuela -- appear to face higher demand risk. This appears to be due to the fact that their exports are disproportionately to the United States, rendering them vulnerable to US crises.

@@

As discussed above, the model accepts several input parameters, most of which are calibrated from data. Given that this calibration is only approximate, it is important to conduct a sensitivity analysis in order to identify whether the results are extremely sensitive to the input parameters -- the more sensitive they are, the more the model will be biased by missing or incomplete data, and the less reliable we expect it to be. The space of parameters and outputs to search is extremely large; for the purpose of this analysis, I will limit myself to two parameters: the exponential coefficient on the conflict length power law (calibrated from data, as described above) and the default spare production capacity, estimated from SME input. The output I examine is the Overall Ratio, as it provides a global summary of the model condition.

In order to conduct the sensitivity analysis, I iterate over conflict coefficient values between -1.05 and -2.0, and default spare production capacities between 0 and 0.5, both in steps of 0.05. For each combination, I run the model 40 times, 10 iterations each for each combination of Contagion and Assistance, for a total of 2,200 model runs.
